\chapter{Background}
This chapter serves as a fundamental cornerstone in the overall structure of this thesis, as it is designed to provide a comprehensive and in-depth exploration of several key areas of study that are crucial to the understanding of the research presented in the subsequent chapters. Through a detailed exposition, this chapter illuminates the foundational principles and contextual factors that underlie the entire thesis.

The first component of this chapter is dedicated to elucidating the intricate domain of information retrieval, a field that is pivotal in the context of this research. Information retrieval is the process of obtaining relevant information from vast collections of data, and it encompasses various techniques and methodologies aimed at achieving this objective. By delving into this subject matter, the reader is equipped with the essential knowledge to appreciate the challenges and complexities of the thesis's research focus.

Large language models, which are a central focus of this work, are introduced and discussed in this chapter. These models represent a breakthrough in natural language processing and artificial intelligence, capable of comprehending and generating human-like text at an unprecedented scale. Their significance lies in their potential to enhance information retrieval systems, an integral aspect of this research. The chapter expounds on the architecture, capabilities, and potential applications of these models, laying the groundwork for the subsequent exploration of their role in the thesis.

The evaluation of retrieval pipelines, a critical research component, is also thoroughly examined in this chapter. It elucidates the methodologies and metrics used to assess the performance and effectiveness of information retrieval systems. An understanding of evaluation techniques is vital for comprehending the empirical aspects of the research and how improvements in retrieval pipelines are quantified and validated.

Moreover, this chapter introduces the concept of query variations, a pivotal notion that forms the bedrock of subsequent research endeavours. Query variations involve the diversification and adaptation of search queries to enhance the quality and relevance of retrieved information. The chapter explores the rationale behind this concept, its significance, and its potential impact on information retrieval systems, setting the stage for the research that follows.

\input{2Background/Background/informationRetrieval}
\input{2Background/Background/naturalLanguageProcessing}
\input{2Background/Background/effectivenessEvaluation}
\input{2Background/Background/queryVariation}