\chapter{Abstract}
This thesis delves into a comprehensive investigation of the profound significance of query variations in information retrieval pipelines, aiming to replicate and expand upon the seminal research conducted by Gustavo Penha, Arthur Câmara, and Claudia Hauf. The central objective of this study is to delve into how the utilization of query variations, subtly modified versions of search queries, can exert a discernible influence on retrieval efficiency.

The research methodology adopted in this study involved the application of ten well-established methods for generating query variations. These query variations were systematically introduced into three datasets: ANTIQUE, TREC-DL-2019, and DL-TYPO. This rigorous process revealed some intriguing findings, notably that methods involving synonyms and paraphrasing often resulted in queries with nuanced meanings, primarily due to the compounded effect of misspellings within the queries.

The experimental design encompassed the utilization of BM25 as an initial retrieval model, followed by the re-ranking of the top 100 retrieved results using various models. This procedure was meticulously executed for both the original search queries and their respective variations, with the evaluation of retrieval effectiveness being quantified through the nDCG@10 metric. Additionally, the study examined the impact of combining rankings from different query variation categories, providing a comprehensive assessment of the potential benefits of cross-category query combinations.

The outcomes of the experiments unequivocally showcased that query variations indeed possess a profound influence on retrieval effectiveness. The replication phase of the study effectively reaffirmed the importance of investigating query variations while introducing the novel DL-TYPO dataset. This phase also underscored the necessity of considering dataset-specific characteristics when assessing the impact of query variations. Notably, it was found that 51, 41, and 32 instances in ANTIQUE, TREC-DL-2019, and DL-TYPO, respectively, exhibited significant declines in retrieval effectiveness when query variations were introduced. Although combining query variation categories often yielded improvements in results, it generally fell short of matching the effectiveness of the original, unaltered queries.

In summation, this thesis shows comprehension of the profound influence that query variations exert on the robustness of retrieval pipelines, reaffirming their pivotal role in evaluating retrieval systems. The study successfully replicated and expanded upon the initial findings, offering valuable insights into the impact of different query variation generators. For those interested in further exploration and replication, the code and associated materials are conveniently accessible via the following link: \url{https://github.com/krista-bradshaw/query_variation_generators}.