\chapter{Results \& Discussion}
This chapter offers a detailed and thorough account of the research outcomes, emphasising their significance and the invaluable insights they bring to the forefront. The results and discussions have been meticulously structured into three key sections to maintain clarity and facilitate a comprehensive understanding of the research findings while adhering to the original formatting.

The initial section of this chapter is dedicated to the critical task of generating various query variations and systematically evaluating their validity. This is a crucial step in the research as it directly addresses our primary research questions. Crafting and assessing these query variations sheds light on the intricate methodology behind them.

The subsequent two sections are categorised in alignment with the research questions they aim to answer. The second section revolves around an in-depth analysis of the primary experiments conducted in the original study, specifically concerning the robustness of the retrieval pipelines in the face of these query variations. The outcomes and implications of reproducing the original study's main experiments are scrutinised, offering a comprehensive evaluation of the retrieval pipelines' resilience to these query variations.

The third section is dedicated to the exploration of the second research question. Here, investigation is extended by delving into the effects of query variations on an additional dataset known as DL-TYPO. This in-depth analysis allows us to draw insightful comparisons and contrasts, shedding light on how query variations impact different datasets, thus broadening the scope of our research and offering a richer perspective.

\input{5Results/method-quality}
\input{5Results/main/robust.tex}
\input{5Results/fusion/fusion.tex}
